---
title: "EQ2 Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EQ2 Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The EQ2 package imports data from the National Oceanic and Atmospheric Administration which includes several thousands of years of earthquake data from around the world.

## Installation

You can install this package from github using the following command:

```{r, eval=FALSE}
library(EQ2)
```

## EQ2 functions:
  
* The functions `eq_clean_data` and `eq_location_clean` will clean the data and return a data.frame with the cleaned data.
* The function eq_map displayes an interactive map of earthquakes around the world, based on parameters of country and/or year.
* The function eq_create_label creates the HTML label that shows up when you hover your mouse over that interactive map.
* The geom geom_timline visualizes the magnitude of earthquakes within a given date range
* The geom geom_timelinelabel annotates the earthquake data
     
## Examples

```{r eval = FALSE}
file_name <- system.file("extdata", "earthquake.txt", package="EQ2")
library(readr)
data <- readr::read_delim(file_name, delim = "\t")
```

Clean the data with `eq_clean_data` and `eq_location_clean`:

```{r eval = FALSE}
clean_data <- eq_clean_data(data)
clean_data <- eq_location_clean(clean_data)
```

To map the earthquakes epicenters in Japan since 2000, use the `eq_map` function:

```{r eval = FALSE}
eq_data <- clean_data                                                     %>%
            dplyr::filter(COUNTRY == "JAPAN" & lubridate::year(DATE) >= 2000) %>%
            eq_map(annot_col = "popup_text")
```

To have data about the earthquake pop-up when you hover your mouse over a radius, use the `eq_create_label`:

```{r eval = FALSE}
eq_data <- clean_data                                                     %>%
            dplyr::filter(COUNTRY == "JAPAN" & lubridate::year(DATE) >= 2000) %>%
            dplyr::mutate(popup_text = eq_create_label(.))                    %>%
            eq_map(annot_col = "popup_text")
```

To show dates, magnitude, and number of deaths based on country and year, use the geom `geom_timeline` and the ggplot function:

```{r eval = FALSE}
eq_data <- clean_data                    %>%                                 
  dplyr::filter(COUNTRY == "JAPAN" & lubridate::year(DATE) >= 2000) %>%
  ggplot(aes(x=DATE,y=COUNTRY,color=TOTAL_DEATHS,size=EQ_PRIMARY)) +
  geom_timeline(alpha=.5) + 
  theme(legend.position="bottom", legend.box="horizontal",
  plot.title=element_text(hjust=0.5)) +
  ggtitle("Earthquakes Visualization Tool") +
  labs(size = "Richter scale value", color = "# deaths")
```

To add annotations to the earthquake data, based on year, use the geom`geom_timelinelabel`:

```{r eval = FALSE}
eq_data <- clean_data                    %>%                                 
  dplyr::filter(COUNTRY == "JAPAN" & lubridate::year(DATE) >= 2000) %>%
  ggplot(aes(x=DATE,y=COUNTRY,color=TOTAL_DEATHS,size=EQ_PRIMARY)) +
  geom_timeline(alpha=.5) +
  geom_timelinelabel(aes(label=LOCATION_NAME),n_max=3) +
  theme(legend.position="bottom", legend.box="horizontal", plot.title=element_text(hjust=0.5)) +
  ggtitle("Earthquakes Visualization Tool") +
  labs(size = "Richter scale value", color = "# deaths")
```

